Numerical simulation of bubble deformation \\ and breakup under simple linear shear flows

### 2023-12-31 Reply-Reviewer2.tex

Response: \\
We agree with the reviewer's opinion. We have changed the ``interface tracking''
wording to ``interface capturing method''.
\\
Please see the 1st line from the top in ``I\hspace{-1.2pt}I\hspace{-1.2pt}I. A. Numerical method and governing equations''.
\\

%%2
\par\noindent
\item
\textsf
{Page 3: The definitions of the dimensionless numbers, Re, and Ca, are not
appropriate.  These numbers are defined here using the properties of air (the
dispersed material), whereas in the related references, which are used for
validation, they are defined using the properties of the continuous phase.
This is not so important in the validation tests where viscosity and density
ratios between the two phases are set equal to one, but it is important for the
new results. In general, the deformation of the inclusion is caused by the
forces exerted on it by the fluid in the matrix, and even more so when the
inclusion is just an air bubble, which has negligible density and viscosity.
Moreover, the capillary number is the ratio of viscous forces over capillarity
and the Reynolds number is the ratio of inertia over viscous forces, and the
viscous forces are negligible for air in comparison to viscous forces in the
liquid.}
\vspace{3 mm}

Response: \\
Thank you very much for your suggestion.  We have modified the definitions of
$Re$ and $Ca$ because we had typos in the definitions of $Re$ and $Ca$.  The
computations presented were performed using $Re$ and $Ca$ defined by the
density and the viscosity of the surrounding liquid.
\\
Please see Eq.(2).
\\

%%3
\par\noindent
\item
\textsf
{Page 3: The direction of gravity is not specified. Is it acting along the
z-direction? If so, see point \#9 below.  If not, in what direction does it
act? In other studies, an air bubble was considered under the action of both
gravity and velocity gradient in the same direction, ref. [3] and [4].
Interesting variations arise from the occurrence of a lift force on a
deformable bubble. Again, these additional works need to be cited and
discussed.}
\vspace{3 mm}

Response: \\
We have added the direction of gravity in Fig. 1 ($z$ direction);
albeit, we prescribe zero gravity in our simulations.  The rationale
for setting gravity equal to zero is given in more detail in our response
to item 9.  Essentially the purpose of this study is to do a controlled
study of the effects of the density and viscosity ratios.  We can isolate
the dependencies on these two ratios by prescribing zero gravity.  We now
reference [3] and [4].
\\

%%4
\par\noindent
\item
\textsf
{Page 4, top: “The singular Heaviside gradient term in the right hand side of
equation (5) is a body force...”. Although this term arises along with gravity
as a body force, in fact, it presumably acts only on the fluid/fluid
interface.}
\vspace{3 mm}

@@

Response: \\
{\color{red} 
%As a quantitative comparison, we have newly made a comparison for bubble 
%deformation as Table I\hspace{-1.2pt}I:
We added a new quantitative comparison for bubble deformation in Table
I\hspace{-1.2pt}I to verify the validity of our computational results; in the
table we compare against the experimental results of Muller-Fischer et al.
(Experiments in fluids, 2008) and Rust and Manga (Journal of Colloid and
Interface Science, 2002). The new numerical results for bubble deformation
indicate that our computations are consistent with past studies.
%We believe that the effectivity of our computations could be proved for bubble deformation.
}
\\

%%9
\par\noindent
\item
\textsf
{The following is the most disturbing issue with this study: It is fine to
examine conditions leading to bubble breakup or other flow instabilities, but
the range of parameter values investigated must correspond to existing
materials and achievable flow conditions.  The authors have given the governing
equations in dimensional form and defined dimensionless numbers incorrectly as
stated in point \#2, but also assigned them values that may not be physically
relevant. Moreover, the bubble size is nowhere stated.................\\
Finally, the Reynolds number of the channel in this case would be: $Re_{\rm
channel} \simeq 6000$.  This high value of Re corresponds to turbulence
conditions.................}
\vspace{3 mm}

Response: \\
As we have stated in response to your comment \#2, correct dimensionless
numbers based on the density and the viscosity of the surrounding liquid were
used in real computations.  Regarding physical properties and conditions in our
computations, we will give you an example for the case of $Ca$ = 0.3 and $Re$
=93.\\

Physical properties (a silicone oil has a very similar physical properties with these values.):\\
\quad $\rho_{\rm m}$ (density of liquid) = 1000 kg/m$^3$, $\mu_{\rm m}$ (viscosity of liquid) = 0.2 Pa$\cdot$s,
$\sigma$ (surface tension) = 0.025 N/m, $R$ (bubble radius) = 5.0 mm.\\
From ${\it \Gamma}$ = $2V/H$ = $2V/6R$, \  $V$ = 3${\it \Gamma} R$ = $\frac{3\sigma Ca}{\mu_{\rm m}} $ = 1.1 m/s.
Also, we obtain $Fr$ (Froude number) = $\frac{{\it \Gamma}R}{\sqrt {gR}}$ = 1.7. \\
In this flow system, the effect of gravity may not be completely negligible. However, in our computations, 
we set $g = 0$ because we wanted to clearly isolate only the effects of 
the density and viscosity ratios. 
From another perspective, it can be said that we investigated 
phenomena in space.
At the end of the Problem Description section, we added a detailed explanation about physical properties and conditions.

### 2023-12-31 Bubble-deformation.tex
paper mods

%
% Note: There were missing references in the .bib file for the following cite tags: 
%       WanMacReni02, OlaSinSak09, ZhaBag11, BayMor11, and Ren08-1. 
%       (Edwin, April 29, 2022)
%
The study of the deformation and breakup of a drop in immiscible viscous
liquids undergoing simple linear shear flow has been investigated extensively
due to its fundamental importance in connection with emulsion and materials
processing, mixing and reaction devices. The pioneering experimental work of
this problem was performed by Taylor in the early 1930s \cite{Tay32, Tay34},
and the subsequent theoretical and experimental progress up to the 1980s and
1990s was reviewed in \cite{Ral84} and \cite{Sto94}, respectively.  By the
2000s, progress in computational fluid dynamics (CFD) techniques and increased
access to powerful computing resources led to a surge of research focused on
direct simulations of this problem.  In particular, detailed computational
investigations of drop breakup, based on a Volume-of-Fluid (VOF) method
\cite{HirNic81} were presented in \cite{LiRenRen00, RenCri01-1, RenCri01-2,
RenCriLi02,KhiRenCri03,Ren06,Ren07,Ren08-2}.  Since then, the literature on
computational studies on the deformation and breakup of a single or several
drops in shear flow has continued to grow \cite{CriGuiAlfBlaLoe03,
InaTomOgi03,ZhaMikBan06, BazAndMei06, JanAnd08,
CroGriSch10,KomShaEskDer14,KomShaEskDer15, IoaLiuZha16,
HerRan17,AmaBalCasOli19, ZhaShuGuaYan21} and a variety of numerical techniques
have been developed to tackle this problem, including boundary-integral
approaches \cite{CriBlaLoe01, JanAnd07}, lattice Boltzmann methods \cite{Ina06,
KomShaEskDer14}, front tracking schemes \cite{UnvTry92}, and {\color{red}
interface-capturing} level set methods \cite{SusSmeOsh94}.  {\color{red} In
contrast, only a few computational studies on bubble deformation and breakup in
shear flow can be found \cite{WeiQiaXu12,WanShiZha15}.  In these previous
studies, the dynamics of bubble deformation in shear flow has mainly been
examined.} {\color{blue} Concerning bubble breakup, Wei et
al.~\cite{WeiQiaXu12} presented only one numerical result for a bubble breakup
process under the condition of $Ca$ (Capillary number) = 35.}
% {\color{blue} What distinguishes our work from that in
% \cite{WeiQiaXu12,WanShiZha15}?}
This type of bubble dynamics in shear flow is of critical importance for a
variety of scientific and engineering processes. We refer the reader to the
following experimental studies relating to bubble deformation in foaming
processes, microfluidic devices, and micro bubbles in the blood circulation
system,\cite{ChuFinBouAtaHamPug19,MulTobDreFisWin08,BenRodFauPinFerPerGarMirLim18,DreSai15}.
In particular, it is the study of bubble deformation as it pertains to high
performance plastics applications which motivates this work.  In this article,
we present computational studies of shear-driven deformation and breakup of a
bubble in insoluble viscous liquids.  Studying bubble break-up via computation
rather than experiments simplifies the process of setting a combination of
precise simple shear flow conditions, low $Ca$ conditions, low density ratio,
and low viscosity ratios.  The physical properties that distinguish bubble and
drop studies are expressed in terms of the density ratio $\lambda = \rho_{\rm
b} / \rho_{\rm m}$ and the viscosity ratio $\eta = \mu_{\rm b} / \mu_{\rm m}$,
where $\rho$ is the fluid density, $\mu$ is the viscosity and the subscripts
``b'' and ``m'' denote the ``bubble'' or ``drop'' and the ``matrix fluid'',
respectively.  For a bubble in an insoluble viscous liquid, $\lambda \simeq 0$
and $\eta \simeq 0$, whereas most studies dealing with a drop in an immiscible
viscous liquid take $\lambda =1$ and $\eta \simeq 1$.

@@

When comparing with previous drop studies, we fix $\lambda=1$.  As a result,
(for $\lambda=1$) the following dimensionless physical parameters are used to
describe the problem of drop deformation/breakup in shear flow
%
\begin{eqnarray}\label{dimensonless}
  Re = \frac{\rho_{\rm {\color{red}m}}UR}{\mu_{\rm {\color{red}m}}}, \quad
  Ca = \frac{\mu_{\rm {\color{red}m}}U}{\sigma}, \quad
  \eta = \frac{\mu_{\rm b}}{\mu_{\rm m}}
\end{eqnarray}
%
$U$ is the velocity scale and $\sigma$ denotes the surface tension.  For the
problem of shear-induced drop deformation and breakup, the velocity is set to
$U = \mathit{\Gamma} R$, where the shear-rate is $\mathit{\Gamma} = 2V/H$.  As
mentioned in the introduction, most previous drop studies set $\eta =
1$\cite{?}.  Thus, for comparison with previous drop deformation and breakup
problems, we set $\lambda = \eta = 1$ (and also neglect the effect of gravity
so that $g=0$). On the other hand, in our computations for bubble deformation,
we set the density and viscosity of air to be $\rho_{\rm b}$ = 1.2 kg/m$^{3}$
and $\mu_{\rm b} = 1.8 \times 10^{-5}$ Pa$\cdot$s respectively.  We emphasize
that for consistency with previous studies (\cite{LiRenRen00, RusMan02,
MulTobDreFisWin08, KomShaEskDer14, AmaBalCasOli19}), 
%{\color{blue} cite some of these previous studies which examine deformation
%and breakup of a bubble or drop as a function of the $Re$ and $Ca$ numbers.}
we computationally examine the deformation and breakup of a bubble in simple
linear shear flow as a function of the $Re$ and $Ca$ numbers.  {\color{red}
That is to say, by setting $g=0$, we are isolating the effect of only varying
$Re$ and $Ca$ on bubble deformation and breakup.  In our controlled study, we
determine the critical $Re_{\rm c}$ versus $Ca$ curve in which $Re_{\rm c}$
corresponds to the threshold of bubble (or drop) breakup.  We determine the
critical $Re_{\rm c}$ versus $Ca$ curve for strategic pairs of the density
ratio and viscosity ratio. }

@@

{\color{red}
The temporal discretization of our numerical method is an operator split
projection method as described by \cite{SUSSMAN2007469}.  An outline of 
our method is as follows (see \cite{SUSSMAN2007469}, section 4, for more 
details): }
\begin{description}
\item[Step 1. CLSVOF interface advection]
 \begin{eqnarray*}
 \phi^{n+1}=\phi^{n}-\Delta t[\bmu\cdot\nabla\phi]^{n} \\
 F^{n+1}=F^{n}-\Delta t[\bmu\cdot\nabla F]^{n} 
 \end{eqnarray*}
\item[Step 2. Nonlinear advective force terms]
 \begin{eqnarray*}
 {\cal A}=[\bmu\cdot\nabla\bmu]^{n}
 \end{eqnarray*}
\item[Step 3. Viscosity force]
 \begin{eqnarray*}
 \frac{\bmu^{\ast}-\bmu^{n}}{\Delta t}=-{\cal A}+g\vec{z}-
  [\nabla p/\rho]^{n}+\frac{1}{2}\frac{
    \nabla\cdot(2\mu\bmD^{n})+ 
    \nabla\cdot(2\mu\bmD^{\ast})}{\rho}
 \end{eqnarray*}
\item[Step 4. Pressure projection]
 \begin{eqnarray*}
 \bmV=\bmu^{n}+\Delta t(-{\cal A}+g\vec{z}+
   \frac{1}{2}\frac{
    \nabla\cdot(2\mu\bmD^{n})+ 
    \nabla\cdot(2\mu\bmD^{\ast})}{\rho})
 \end{eqnarray*}
 \begin{eqnarray*}
 \nabla\cdot\frac{\nabla p}{\rho}=\frac{1}{\Delta t}\nabla\cdot\bmV
 \end{eqnarray*}
 \begin{eqnarray*}
 \bmu^{n+1}=\bmV-\Delta t \frac{\nabla p}{\rho}
 \end{eqnarray*}
\end{description}

>>>>

%
% Note: There were missing references in the .bib file for the following cite tags: 
%       WanMacReni02, OlaSinSak09, ZhaBag11, BayMor11, and Ren08-1. 
%       (Edwin, April 29, 2022)
%
The study of the deformation {\em and breakup} of a drop in immiscible viscous
liquids undergoing simple linear shear flow has been investigated extensively
due to its fundamental importance in connection with emulsion and materials
processing, mixing and reaction devices. The pioneering experimental work of
this problem was performed by Taylor in the early 1930s \cite{Tay32, Tay34},
and the subsequent theoretical and experimental progress up to the 1980s and
1990s was reviewed in \cite{Ral84} and \cite{Sto94}, respectively.  By the
2000s, progress in computational fluid dynamics (CFD) techniques and increased
access to powerful computing resources led to a surge of research focused on
direct simulations of this problem.  In particular, detailed computational
investigations of drop breakup, based on a Volume-of-Fluid (VOF) method
\cite{HirNic81} were presented in \cite{LiRenRen00, RenCri01-1, RenCri01-2,
RenCriLi02,KhiRenCri03,Ren06,Ren07,Ren08-2}.  Since then, the literature on
computational studies on the deformation and breakup of a single or several
drops in shear flow has continued to grow \cite{CriGuiAlfBlaLoe03,
InaTomOgi03,ZhaMikBan06, BazAndMei06, JanAnd08,
CroGriSch10,KomShaEskDer14,KomShaEskDer15, IoaLiuZha16,
HerRan17,AmaBalCasOli19, ZhaShuGuaYan21} and a variety of numerical techniques
have been developed to tackle this problem, including boundary-integral
approaches \cite{CriBlaLoe01, JanAnd07}, lattice Boltzmann methods \cite{Ina06,
KomShaEskDer14}, front tracking schemes \cite{UnvTry92}, and {\color{red}
interface-capturing} level set methods \cite{SusSmeOsh94}.  {\color{red} 
In contrast, few computational studies on bubble 
deformation {\em and breakup} can be found.  
We remark that there has been a 
number of articles on the study of lift of {\em slightly}
deformable bubbles\cite{ErvinANDTryggvason1997,legendre1998lift}.
We reiterate though, that for bubble deformation {\em and breakup} in
shear flows, only a few articles exist: \cite{WeiQiaXu12,WanShiZha15}.  
In these previous
studies, the dynamics (e.g. rotation angle) of bubble deformation 
in shear flow was mainly examined.} 
{\color{blue} Concerning bubble breakup, Wei et
al.~\cite{WeiQiaXu12} presented only one numerical result for a bubble breakup
process under the condition of $Ca$ (Capillary number) = 35.}
% {\color{blue} What distinguishes our work from that in
% \cite{WeiQiaXu12,WanShiZha15}?}
Bubble dynamics in shear flow, {\em including breakup},
is of critical importance for a
variety of scientific and engineering processes. We refer the reader to the
following experimental studies relating to bubble deformation in foaming
processes, microfluidic devices, and micro bubbles in the blood circulation
system,\cite{ChuFinBouAtaHamPug19,MulTobDreFisWin08,BenRodFauPinFerPerGarMirLim18,DreSai15}.
In particular, it is the study of bubble deformation as it pertains to high
performance plastics applications which motivates this work.  In this article,
we present computational studies of shear-driven deformation 
{\em and breakup} of a
bubble in insoluble viscous liquids.  Studying bubble break-up via computation
rather than experiments simplifies the process of setting a combination of
precise simple shear flow conditions, low $Ca$ conditions, low density ratio,
and low viscosity ratios.  The physical properties that distinguish bubble and
drop studies are expressed in terms of the density ratio $\lambda = \rho_{\rm
b} / \rho_{\rm m}$ and the viscosity ratio $\eta = \mu_{\rm b} / \mu_{\rm m}$,
where $\rho$ is the fluid density, $\mu$ is the viscosity and the subscripts
``b'' and ``m'' denote the ``bubble'' or ``drop'' and the ``matrix fluid'',
respectively.  For a bubble in an insoluble viscous liquid, $\lambda \simeq 0$
and $\eta \simeq 0$, whereas most studies dealing with a drop in an immiscible
viscous liquid take $\lambda =1$ and $\eta \simeq 1$.

@@

When comparing with previous drop studies, we fix $\lambda=1$.  As a result,
(for $\lambda=1$) the following dimensionless physical parameters are used to
describe the problem of drop deformation/breakup in shear flow
%
\begin{eqnarray}\label{dimensonless}
  Re = \frac{\rho_{\rm {\color{red}m}}UR}{\mu_{\rm {\color{red}m}}}, \quad
  Ca = \frac{\mu_{\rm {\color{red}m}}U}{\sigma}, \quad
  \eta = \frac{\mu_{\rm b}}{\mu_{\rm m}}
\end{eqnarray}
%
$U$ is the velocity scale and $\sigma$ denotes the surface tension.  For the
problem of shear-induced drop deformation and breakup, the velocity is set to
$U = \mathit{\Gamma} R$, where the shear-rate is $\mathit{\Gamma} = 2V/H$.  As
mentioned in the introduction, most previous drop studies set $\eta =
1$ (e.g. \cite{LiRenRen00}).  
Thus, for comparison with previous drop deformation and breakup
problems, we set $\lambda = \eta = 1$ (and also neglect the effect of gravity
so that $g=0$). On the other hand, in our computations for bubble deformation,
we set the density and viscosity of air to be $\rho_{\rm b}$ = 1.2 kg/m$^{3}$
and $\mu_{\rm b} = 1.8 \times 10^{-5}$ Pa$\cdot$s respectively.  We emphasize
that for consistency with previous studies (\cite{LiRenRen00, RusMan02,
MulTobDreFisWin08, KomShaEskDer14, AmaBalCasOli19}), 
%{\color{blue} cite some of these previous studies which examine deformation
%and breakup of a bubble or drop as a function of the $Re$ and $Ca$ numbers.}
we computationally examine the deformation and breakup of a bubble in simple
linear shear flow as a function of the $Re$ and $Ca$ numbers.  {\color{red}
That is to say, by setting $g=0$, we are isolating the effect of only varying
$Re$ and $Ca$ on bubble deformation and breakup.  In our controlled study, we
determine the critical $Re_{\rm c}$ versus $Ca$ curve in which $Re_{\rm c}$
corresponds to the threshold of bubble (or drop) breakup.  We determine the
critical $Re_{\rm c}$ versus $Ca$ curve for strategic pairs of the density
ratio and viscosity ratio. }

@@

{\color{red}
The temporal discretization of our numerical method is an operator split
projection method as described by \cite{SUSSMAN2007469}.  An outline of 
our method is as follows (see \cite{SUSSMAN2007469}, section 4, for more 
details): }
\begin{description}
\item[Step 1. CLSVOF interface advection]
 \begin{eqnarray*}
 \phi^{n+1}=\phi^{n}-\Delta t[\bmu\cdot\nabla\phi]^{n} \\
 F^{n+1}=F^{n}-\Delta t[\bmu\cdot\nabla F]^{n} 
 \end{eqnarray*}
\item[Step 2. Nonlinear advective force terms]
 \begin{eqnarray*}
 {\cal A}=[\bmu\cdot\nabla\bmu]^{n}
 \end{eqnarray*}
\item[Step 3. Viscosity force]
 \begin{eqnarray*}
 \frac{\bmu^{\ast}-\bmu^{n}}{\Delta t}=-{\cal A}+g\vec{z}-
  [\nabla p/\rho]^{n}+\frac{1}{2}\frac{
    \nabla\cdot(2\mu\bmD^{n})+ 
    \nabla\cdot(2\mu\bmD^{\ast})}{\rho}
 \end{eqnarray*}
\item[Step 4. Pressure projection and ghost fluid surface tension algorithm]
 \begin{eqnarray*}
 \bmV=\bmu^{n}+\Delta t(-{\cal A}+g\vec{z}+
   \frac{1}{2}\frac{
    \nabla\cdot(2\mu\bmD^{n})+ 
    \nabla\cdot(2\mu\bmD^{\ast})}{\rho})
 \end{eqnarray*}
 \begin{eqnarray*}
 \bmV=\bmV-
   \Delta t \frac{\sigma\kappa(\phi^{n+1})}{\rho}\nabla \Hea(\phi^{n+1}) 
 \end{eqnarray*}
 \begin{eqnarray*}
 \nabla\cdot\frac{\nabla p}{\rho}=\frac{1}{\Delta t}\nabla\cdot\bmV
 \end{eqnarray*}
 \begin{eqnarray*}
 \bmu^{n+1}=\bmV-\Delta t \frac{\nabla p}{\rho}
 \end{eqnarray*}
\end{description}
